---
title: "Telecom Customer Churn Prediction"
author: "Goutham Polamreddy"
output: 
  pdf_document: 
    fig_height: 4
    fig_width: 4
    number_sections: yes
    toc: yes
    toc_depth: 5
---

# Environment setup

## Setting up working directory

```{r}
setwd("C:\\Users\\Goutham Reddy\\Desktop\\R files\\exploRe")
getwd()
```

## Loading required libraries

```{r Libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(corrplot)
library(ggplot2)
library(Amelia)
library(mlbench)
library(rms)
library(pROC)
library(e1071)
library(caTools)
library(class)
library(ROCR)
library(ineq)
library(InformationValue)
```

# Exploratory Data Analysis

## Importing data

* Importing data using readxl package

```{r}
customerChurnData = read_excel("Cellphone.xlsx", sheet = 2)

# copy data frame in case features in original form are required in analysis
customerChurnDataCopy = customerChurnData
```

## Summary and Data preparation

* summary shows it is an unbalanced data set, so while evaluation model performance it is recommended to check measures like specificity/Sensitivity.


```{r}
summary(customerChurnData)

```

* Converting appropriate variables to factors help in executing models.

```{r}
customerChurnData$Churn = as.factor(customerChurnData$Churn)
customerChurnData$DataPlan = as.factor(customerChurnData$DataPlan)
customerChurnData$ContractRenewal = as.factor(customerChurnData$ContractRenewal)
```

## Univariate and Bi-variate analysis

### Univariate Plots and analysis

* Below showing boxplots of continuous variables and their outliers

```{r}
par(mfrow=c(1,4))

for(i in 5:9) {
    boxplot(customerChurnData[,i], main=names(customerChurnData)[i])
}

for(i in 9:11) {
    boxplot(customerChurnData[,i], main=names(customerChurnData)[i])
}

```

### Bivariate Plots

* Quite evident from below data usage between users with and without data plan

```{r}
qplot(customerChurnData$DataPlan, customerChurnData$DataUsage, geom = "boxplot")
```

* From the plot below, people who opt out relatively uses less data

```{r}
qplot(customerChurnData$Churn, customerChurnData$DataUsage, geom = "boxplot")
```

* Other factors like Roaming minutes and Overage fee is similar for both users who opt out and ones who continue

## Multicollinearity check

* Dataplan and data usage for eg are highly correlated, these correlated pairs are tackled while computing the model itself.

```{r}
churnCorrMatrix = cor(customerChurnData[,-c(1,3,4)])

corrplot(churnCorrMatrix, method = "number")

```


## Missing values and outliers

* Clear from below figure that there are no missing values

```{r}
missmap(customerChurnData, col=c("blue", "red"), legend=FALSE)
```

# Model Building 

## Logistic regression

```{r}

churnGLM = glm(Churn ~ AccountWeeks+ContractRenewal+DataPlan+DataUsage+CustServCalls+DayMins    +DayCalls+MonthlyCharge+OverageFee+RoamMins, data = customerChurnData, family = "binomial")

summary(churnGLM)

vif(churnGLM)


#Variables are inflated let's try to remove some variable and see how it resolves it

churnGLM = glm(Churn ~ AccountWeeks+ContractRenewal+DataPlan+CustServCalls+DayMins    +DayCalls+MonthlyCharge+OverageFee+RoamMins, data = customerChurnData, family = "binomial")

vif(churnGLM)


churnGLM = glm(Churn ~ AccountWeeks+ContractRenewal+DataPlan+CustServCalls+
                 DayMins+DayCalls+OverageFee+RoamMins, data = customerChurnData, family = "binomial")

vif(churnGLM)

qplot(customerChurnData$Churn,churnGLM$fitted.values, geom = "boxplot")
```

```{r}
churn.predicted=ifelse(churnGLM$fitted.values<0.15,0,1)
churnConfMat = table(customerChurnData$Churn,churn.predicted)

print(churnConfMat)

churnErrorRate = (churnConfMat[1,2]+ churnConfMat[2,1])/nrow(customerChurnData)

print(churnErrorRate)

roc(customerChurnData$Churn,churnGLM$fitted.values)
plot.roc(customerChurnData$Churn,churnGLM$fitted.values)

```

## Implement Naïve Bayes Classifier

```{r}

churnNBM = naiveBayes(Churn ~ AccountWeeks+ContractRenewal+DataPlan+CustServCalls
                      +DayMins+DayCalls+OverageFee+RoamMins,data=customerChurnData)

print(churnNBM)

churnNBMPredictProb = predict(churnNBM,type="raw",newdata=customerChurnData)


qplot(customerChurnData$Churn,churnNBMPredictProb[,2], geom = "boxplot")

#churnNBMPredict = predict(churnNBM,newdata=customerChurnData)

churnNBMPredict = ifelse(churnNBMPredictProb[,2]<0.125,0,1)

nbmConfMat = table(customerChurnData$Churn, churnNBMPredict)

nbmErrorRate = (nbmConfMat[1,2]+nbmConfMat[2,1])/nrow(customerChurnData)

print(nbmConfMat)
print(nbmErrorRate)
```

## KNN Classifier

* Since Dataset contains both Continuous and categorical features, converting categorical to numerical by considering each factor as a number.

### Data preparation for kNN

```{r}
summary(customerChurnDataCopy)

nrow(customerChurnDataCopy)

```


### Splitting the data test and train

```{r message=FALSE, warning=FALSE}
set.seed(101) 
sample = sample.split(customerChurnDataCopy$Churn, SplitRatio = .70)
trainCustChurnData = subset(customerChurnDataCopy, sample == TRUE)
testCustChurnData  = subset(customerChurnDataCopy, sample == FALSE)

nrow(trainCustChurnData)
nrow(testCustChurnData)

```

### Applying kNN

* since features are in various scales, normalizing them gives better results.
* Here removing Datausage and Monthly charge, since they were causing other features' variance to inflate

```{r}

churnKNNPred = knn(scale(trainCustChurnData[,-c(1)]), 
                   scale(testCustChurnData[,-c(1)]), trainCustChurnData$Churn, k=10)
table(testCustChurnData$Churn, churnKNNPred)
  
```

### Finding out the best possible N

* Since sensitivity is important here, choosing k value which yields the best Sensitivity value below

```{r}

for (i in 10:1) {

  churnKNNPred = knn(scale(trainCustChurnData[,-c(1)]), 
                     scale(testCustChurnData[,-c(1)]), trainCustChurnData$Churn, k=i)

  knnConfMat = table(testCustChurnData$Churn, churnKNNPred)

  sens = knnConfMat[2,2]/(knnConfMat[2,1]+knnConfMat[2,2])

  cat("k=",i,"sensitivity=",sens*100,"%","\n")
    
}

```

* From the above output we can see that k=3 has better sensitivity, while k=1, k=2 might be considering noise.

* So making model with k = 3

```{r}

churnKNNPred = knn(scale(trainCustChurnData[,-c(1)]), 
                   scale(testCustChurnData[,-c(1)]), trainCustChurnData$Churn, k=3)

knnConfMat = table(testCustChurnData$Churn, churnKNNPred)
```


# Evaluation metrics

## Custom Methods for preformance metrics

Making couple of functions to help output some performance metrics by passing 

```{r}
PerfMetrics <- function(prob, tar, cMatrix){
  
  predObj = prediction(prob, tar)
  perf = performance(predObj, "tpr", "fpr")

  #ROC Curve  
  plot(perf)

  #KS 
  KS = max(perf@y.values[[1]]-perf@x.values[[1]])
  
  metrics = list()
  metrics["KS"] = KS
  

  #area under the curve
  auc = performance(predObj,"auc"); 
  auc = as.numeric(auc@y.values)
  
  metrics["auc"] = auc

  #Gini
  gini = ineq(prob, type="Gini")
  metrics["gini"] = gini

  #Concordance ratio
  conc = Concordance(actuals=tar, predictedScores=prob)
  metrics["conc"] = conc
  
  #errorRate
  errorRate = (cMatrix[1,2]+cMatrix[2,1])/length(tar)
  metrics["errorRate"] = errorRate
  
  #Accuracy
  accuracy = 1 - errorRate
  metrics["accuracy"] = accuracy
  
  #Sensitivity
  sensitivityValue = cMatrix[2,2]/(cMatrix[2,1]+cMatrix[2,2])
  metrics["sensitivity"] = sensitivityValue
  
  #Specificity
  specificityValue = cMatrix[1,1]/(cMatrix[1,1]+cMatrix[1,2])
  metrics["specificity"] = specificityValue
  
  return(metrics)

}

confMatMetrics <- function(cMatrix){
  
  metrics = list()
  
  #errorRate
  errorRate = (cMatrix[1,2]+cMatrix[2,1])/(cMatrix[1,1]+cMatrix[1,2]+cMatrix[2,1]+cMatrix[2,2])
  metrics["errorRate"] = errorRate
  
  #Accuracy
  accuracy = 1 - errorRate
  metrics["accuracy"] = accuracy
  
  #Sensitivity
  sensitivityValue = cMatrix[2,2]/(cMatrix[2,1]+cMatrix[2,2])
  metrics["sensitivity"] = sensitivityValue
  
  #Specificity
  specificityValue = cMatrix[1,1]/(cMatrix[1,1]+cMatrix[1,2])
  metrics["specificity"] = specificityValue
  
  return(metrics)
}

```


## Logistic Regression

```{r}
logisticRegression = PerfMetrics(churnGLM$fitted.values, customerChurnData$Churn, churnConfMat)
print(logisticRegression)
```



## Naive Bayes

```{r}
# Naive Bayes performance metrics

naiveBayesMetrics = PerfMetrics(churnNBMPredictProb[,2], customerChurnData$Churn, nbmConfMat)
print(naiveBayesMetrics)
```

## K Nearest Neighbours

```{r}

knnPerfMetrics = confMatMetrics(knnConfMat)

print(knnPerfMetrics)

```

## Comparision

Considering different metrics kNN seems to be the best bet considering an accuracy of 0.888

# Interpretation and Recommendations

Since kNN is found as best for this dataset. Following interpretations and recommendations can be made

* kNN gives predictions by comparing each sample with its nearest neighbours. So by we comparing each customer's behaviour, similar behaviour can be estimated. *Customer profiling can be advantageous*

* Suppose if customers who are levied with roaming charges are chrurning out, then it is possible that a customer who is roaming will likely opt out.